---
output:
params: 
  set_title: "My Title!"
  pair: "this.pair"
  outpath: "my.location"
  bam: "tumor_bam"
fig_crop: false
---

---
title: `r params$set_title`
---
    
```{r, include = FALSE, echo = FALSE, message = TRUE}
library(skitools)
library(skidb)
library(ggplot2)
library(ggpubr)
library(ggforce)
library(ggpubr)
library(png)
library(kableExtra)

message("Loaded Packages")
```

***

# Introduction 

This is a case report based upon whole genome sequencing performed between Weill
Cornell's [Institute for Precision Medicine](https://eipm.weill.cornell.edu/), and the [New York Genome Center](https://http://nygenome.org/), with
analysis performed by members of the [Imielinski Lab](https://www.mskilab.org/).

This report draws on methods developed by many groups and individuals, with
appropriate citations to be found in the footnotes.

The goal of this report is fourfold:

* To provide a sense of WGS library quality
* To frame molecular features of this case in the setting of other WGS datasets
* To explore structural alterations to the genome which cannot be determined
with less-than-whole genome sequencing
	+ Including an analysis of complex structural features
* To report potentially actionable and clinically relevant mutations


***

<div class =  "row">
<div class =  "col-md-6">

# Quality Metrics

<br>

WGS Libary Quality estimates are based upon three parameters:

* Ploidy-corrected sample purity, which reflects an estimate of the fraction of
tumor-derived vs normal-derived nuclear material from which the WGS library was
made.

* Coverage depth: Number of whole-genome equivalents (~3e9 base pairs) returned
by the sequencing experiment

* Coverage variance: The degree of local variability in read coverage, generally
indicative of library preparation quality.  A measure of noise.  High PCR duplicates and (Very)
highly rearranged tumors are two cases where this value will be high. The method
used is Derivative Log Ratio Spread (DLRS)

</div>

<div class =  "col-md-6">
### 

```{r, message=FALSE, echo=FALSE, fig.align = "left", warning=FALSE}
if(file.exists(paste0(params$outpath,"/purity_donut.png"))){
  purity = readPNG(paste0(params$outpath,"/purity_donut.png"))
  try(grid::grid.raster(purity))} else { message("No Purity Donut!")}
```
</div>
</div>

<div class =  "row">

<div class =  "col-md-6">
### Coverage
```{r, message = FALSE, echo = FALSE}

this.file = gsub("([^/]*)$",
                 "aligned.alignment_summary_metrics",
                 normalizePath(params$bam))

if(file.exists(this.file)){y = fread(this.file)
  bases = y[3, "PF_HQ_ALIGNED_BASES"]
  depth = bases/sum(hg_seqlengths())
  ggplot(depth, aes(x = params$pair, y = PF_HQ_ALIGNED_BASES)) +
    geom_col(fill = "slateblue3") +
    geom_text(aes(y = PF_HQ_ALIGNED_BASES + 5, label = floor(PF_HQ_ALIGNED_BASES)), size = 6) +
    ylim(0, 120) +
    theme_classic() +
    labs(fill = "", y = "Genome Equivalents", title = paste0(" "),size = 20) +
    theme(legend.position = "none",
          axis.text.x = element_text(size = 25, angle = 0,vjust = 0.5),
          axis.text.y = element_text(size = 20, angle = 0,vjust = 0),
          axis.title.y = element_text(size = 30),
          axis.title.x = element_blank()
          )
} else{ message("No Picard Alignment Summary!")}
```
</div>

<div class =  "col-md-6">
### Coverage Variance
```{r, message=FALSE, echo=FALSE, fig.align = "left"}
y = read.csv(paste0(params$pair, "_quality_metrics.csv"))
y$pair = params$pair
ref = readRDS("/gpfs/commons/groups/imielinski_lab/DB/modules/CaseReport/db/bulk.stats.rds")

require(gridExtra) 

barplot = ggplot(y, aes(x = pair, y = variance)) +
           geom_col(fill = "slateblue3") +
  ylim(0,0.5) +
  theme_classic() +
  labs(fill = "", y = "DLRS Coverage Variance", title = paste0(" "),size = 20) +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 25, angle = 0,vjust = 0.5),
        axis.text.y = element_text(size = 20, angle = 0,vjust = 0),
        axis.title.y = element_text(size = 30),
        axis.title.x = element_blank()
        )
sinaplot = ggplot(ref, aes(x = "IPM Cohort Data", y = variance)) +
  geom_sina(color = "grey30", size = 3) +
  ylim(0,0.5) +
  labs(fill = " ", y = "", title = paste0(" "), size = 20) +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 25, angle = 0,vjust = 0.5),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank())
grid.arrange(barplot, sinaplot, ncol = 2)

```
</div>
</div>

## 

***

# Mutational Signatures and Contexts

Mutational burden, as measured in units of somatic mutations per megabase,
demonstrates disease specificity in distribution.  Individual tumors, however,
may fall well outside expected values for their type, presumably depending upon underlying
mutational processes.


```{r, message=FALSE, echo=FALSE}
tmb = readPNG(paste0(params$pair,"TMB_Context.png"))
grid::grid.raster(tmb)
```

<br>

SNV Signature analysis can provide a sense of the mutational processes
underlying somatic variants in a tumor.  In this analysis, single nucleotide
variants have been analysed via the deconstructSigs pipeline and the burden of
COSMIC mutational signatures has been estimated.

```{r, message=FALSE, echo=FALSE,fig.asp = 0.50}
sigs = readPNG(paste0("SNV_Signatures.png"))
grid::grid.raster(sigs)
```

***

<div class =  "row">
<div class =  "col-md-6">

SNV counts contributing to the above fractional signatures are reported here.  Values are fractional signature multiplied by total number of filtered and contributing SNVs to arrive at total contributing SNV count.  

<br>


</div>

<div class =  "col-md-6">
### 

```{r, message=FALSE, echo=FALSE, fig.align = "left", warning=FALSE}
if(file.exists(paste0(params$outpath,"/SNV_Sig_burden.png"))){
  sigburden = readPNG(paste0(params$outpath,"/SNV_Sig_burden.png"))
  try(grid::grid.raster(sigburden))} else { message("No signature burden plot!")}
```
</div>
</div>




***

# Driver Events and Actionable Mutations

Multiple databases of actionable events exist.  These are either known drivers,
druggable targets, or variants with prognostic significance.  This report uses
[CIVIC](https://civicdb.org/home) as a resource.  The CIVIC database has been
cross referenced with all of the discovered genomic alterations in this case.

Relevant events are as follows:

```{r, message = FALSE, echo = FALSE}
if(file.exists("CIVIC.hits.rds")){
y = try(readRDS("CIVIC.hits.rds"))
y = y[ ,c(1,4,5,6)]
names(y) = c("Alteration", "gene", "CIVIC ID", "Number of Publications")
y = y[!duplicated(y$Alteration), ]
knitr::kable(y, padding = 200)}
```
<br>

# {.tabset .tabset-fade .tabset-pills}

```{r, results='asis', echo = FALSE}
if(!file.exists("CIVIC.hits.rds")){
  cat("## No Actionable Mutations Detected", '<br>')} else {
for (n in 1:nrow(y)){
  cat("##", y$Alteration[n], '<br>', '\n\n')
  evidence = readRDS("CIVIC.report.rds")
  z = evidence[gene == y$gene[n] , -c(2, 11)]
  wtf = unlist(lapply(1:nrow(z), function(n){
    eh = z$drugs[[n]]$name
    eh = paste(eh, collapse = ", ")
    return(eh)}))
  z$drugs = wtf
  z$phenotypes = NULL
  print(knitr::kable(z, padding = 200)
        %>% kable_styling(fixed_thead = T)
        %>% column_spec(column = 3, width_min = "14cm", include_thead = TRUE)
        %>% scroll_box(width = "100%", height = "10cm")
        )
  cat('\n', '<br>', '\n\n')
}}
```

*** 

# Structural Events {.tabset .tabset-fade .tabset-pills}

## Simple Structural Events {.tabset .tabset-fade .tabset-pills}

```{r, results='asis', echo = FALSE, out.width = "100%", fig.align = "center"}
files <- dir("simple_events/")
headers <- gsub(".*-","",gsub(".png","", dir("simple_events/")))

if(length(headers) >0) {
for (n in 1:length(headers)){
  cat("###", headers[n], '<br>', '\n\n')
  plot.new()
  usr = par(mar=c(0,0,0,0), xpd=NA, mgp=c(0,0,0), oma=c(0,0,0,0), ann=F)
  this.png = try(readPNG(paste0('simple_events/',files[n])))
  try(rasterImage(this.png,usr[1],usr[3],usr[2],usr[4]))
  try(grid::grid.raster(this.png,width = 1.0, just = "center"))
  cat('\n', '<br>', '\n\n')
}}
```


## Chromothripsis Events {.tabset .tabset-fade .tabset-pills}

```{r, results='asis', echo = FALSE, out.width = "200%", fig.align = "left"}
files <- dir("chromothripsis_events/")
headers <- gsub(".*-","",gsub(".png","", dir("chromothripsis_events/")))

if(length(headers) >0) {
for (n in 1:length(headers)){
  cat("###", headers[n], '<br>', '\n')
  plot.new()
  this.png = try(readPNG(paste0("chromothripsis_events/",files[n])))
  try(grid::grid.raster(this.png,width = 1.0))
  cat('\n', '<br>', '\n\n')
}}
```


## BFBs {.tabset .tabset-fade .tabset-pills}
```{r, results='asis', echo = FALSE}
files <- dir("bfb_events/")
headers <- gsub(".*-","",gsub(".png","", dir("bfb_events/")))

if(length(headers) >0) {
for (n in 1:length(headers)){
  cat("###", headers[n], '<br>', '\n')
  plot.new()
  this.png = try(readPNG(paste0("bfb_events/",files[n])))
  try(grid::grid.raster(this.png,width = 1.0))
  cat('\n', '<br>', '\n\n')
}}
```


## Templated Insertion Chains {.tabset .tabset-fade .tabset-pills}

```{r, results='asis', echo = FALSE}
files <- dir("tic_events/")
headers <- gsub(".*-","",gsub(".png","", dir("tic_events/")))

if(length(headers) >0) {
for (n in 1:length(headers)){
  cat("###", headers[n], '<br>', '\n')
  plot.new()
  this.png = try(readPNG(paste0("tic_events/",files[n])))
  try(grid::grid.raster(this.png,width = 1.0))
  cat('\n', '<br>', '\n\n')
}}
```

# Fusions {.tabset .tabset-fade .tabset-pills}

Fusion Events may be relevant as drivers in certain cancer types.  They may also
arise by chance, as the result of structural variants.  Here we report only the
fusion events which are in frame fusions, which link 2 or more genes.

```{r, results='asis', echo = FALSE}
files <- dir("fusions/")
headers <- gsub(".*-","",gsub(".png","", dir("fusions/")))

if(length(headers) >0) {
for (n in 1:length(headers)){
  cat("##", headers[n], '<br>', '\n')
  plot.new()
  this.png = try(readPNG(paste0("fusions/",files[n])))
 try(grid::grid.raster(this.png,width = 1.0))
  cat('\n', '<br>', '\n\n')
}}
```
